### 2016-06-28 store-consistency.tex
Example done.

\section{Data Stores and Consistency}
\label{sec:store-consistency}

The operational semantics of Fig.~\ref{fig:txnimp} allows operations
to witness arbitrary subsets of the global state, effectively mimicing the
behaviour of an eventually consistent ({\sc ec}) data
store\footnote{Eventual consistency guarantees that in the absence of further
  updates, all reads witness the same global state
  \emph{eventually}. In any finite trace, however, there are no
  guarantees on what a read may witness.}. There are, however, data stores,
such as relational databases, that provide stronger consistency
guarantees than {\sc ec}. Like the isolation levels of transactions,
the consistency level of the underlying store also affects the
semantics of a program in non-trivial ways. In this section, we
demonstrate how the semantics of stronger stores with on-demand weak
isolation can be captured in our operational model.

% A non-trivial $\I$ composed of isolation specifications from
% Fig.~\ref{fig:ansi-isolation} induces the machine to provide
% non-trivial isolation guarantees for transactions. However, weak
% isolation levels often only constrain the visibility sets of
% operations by dictating what \emph{not} to see; not what to see.  For
% instance, \iso{Repeatable Read} isolation prohibits operations of a
% transaction from witnessing different states. It, however, does not
% prohibit all operations of a transaction from witnessing an aribitrary
% subset of the global state. Consequently, the machine can remain an
% {\sc ec} store even while providing non-trivial isolation. How then to
% model the semantics of an {\sc sc} store, such as a relational
% database, with variable (weak) isolation?

First, we observe that the semantics of a weakly-consistent data store
can be captured by store-specific consistency constraints, along with
transaction-specific isolation constraints, via the trace invariant
$\I$. In particular, we can split $\I$ into two components: (1).
$\I_s$, the store-specific invariant, and (2). $\I_c$, the
program-specific (or, client-specific) invariant, to capture
consistency and isolation constraints, resp.  $\I$ is now a
conjunction: $\I \,=\, \lambda\E.~\I_s(\E) \wedge \I_c(\E)$ (often
simplified to $\I \,=\, \I_s\wedge\I_c$).  While the program-specific
trace invariant ($\I_c$) remains an invariant regardless of the
specific consistency and visibility features exposed by the underlying
data store, the store specific invariant ($\I_s$) changes from store
to store depending on the consisteny level.  For an eventually
consistent store, $\I_s$ is simply \emph{true}. Stronger stores, such
as those that support strong consistenty, have non-trivial definitions
for $\I_s$.

A strongly-consistent {\sc sc} store guarantees a total order of all
operations w.r.t $\visZ$ consistent with their chronological order. A
straightforward $\I_s$ for this store is the \C{SC} property
formalized below:
\begin{smathpar}
\begin{array}{l}
  \C{sc}(\E) \;=\; \forall\eta_1,\eta_2.~\{\eta_1,\eta_2\}
  \subseteq \E.\A \conj \id(\eta_1) < \id(\eta_2) \\
  \hspace*{2in}\Rightarrow \underE{\eta_1 \visar \eta_2}
\end{array}
\end{smathpar}
Unfortunately, $\I_s=\C{SC}$ conflicts with all isolation
specifications of Fig.~\ref{fig:ansi-isolation}. For instance,
consider a case where $\I_c(\E) \;=\; \forall
T_i.~\underE{\C{RC}(T_i)}$, a constraint that dictates all
transactions execute under \iso{Read Committed} isolation. Imagine a
sample execution where $\eta_1$'s transaction is not yet committed
when $\eta_2$ is generated. Letting $\eta_1$ be visible to $\eta_2$
violates $\I_c$, whereas not letting it be visible violates $\I_{s}$.
The only way to satisfy both invariants is to rule out all 
executions that interleave the operations of one transaction with the
other, thereby enforcing serializability.  In general, when $\I_s$ conflicts (but is not
inconsistent) with $\I_c$, the only way to enforce both invariant sets
is to restrict concurrency. Clearly, this is unacceptable since it
defeats the very purpose of supporting weak isolation. 
% How then do we enforce weak isolation on a strongly consistent
% machine?

In practice, relational databases resolve such conflicts by
prioritizing weak isolation (thus, concurrency and performance) over
strong consistency, so the execution traces do not necessarily satisfy
{\sc sc}. In particular, visibility constraints imposed by {\sc sc}
are violated iff they are found to be in conflict with the constraints
imposed by a transaction isolation level. In the context of
the aforementioned example, $\eta_1$ is not made visible to $\eta_2$
because doing so would violate $\I_c$. However, if $\I_c$ is true,
then the store makes $\eta_1$ visible to $\eta_2$ to honor its
consistency commitment\footnote{The term \emph{recency
    commitment}~\cite{bailishat} is often used in practice
  to capture the best-effort nature of {\sc sc}.}. We generalize this
approach to any $\I_s$ and $\I_c$ by defining a \emph{maximum
  visiblity principle} to determine an acceptable weakening of $\I_s$
in case of a conflict with $\I_c$.  The principle requires the
weakened consistency guarantee ($\I_s'$) of the store to enforce all
visibility relationships imposed by the actual consistency guarantee
($\I_s$), unless enforcing such a relationship violates $\I_c$.
% The formal definition of the principle is relegated
% to the supplementary in the interest of space, but it can be
% understood in the context of an {\sc sc} store, where it weakens the
% store invariant to the following:
Formally:
\begin{definition}
$\I_s' : \E \rightarrow \Prop$ is said to be a maximum visibility
weakening of $\I_s : \E \rightarrow \Prop$ if and only if:
\begin{itemize}
  \item $\I_s'$ is weaker than $\I_s$: 
      $\forall\E.~ \I_s(\E) \Rightarrow \I_s'(\E)$, and
  \item In every trace $\E$ that satisfies $\I_c$, and for every pair
  of effects $\eta_1$ and $\eta_2$ in $\E$, if $\I_s(\E)$ requires
  $\eta_1$ to be visible to $\eta_2$, then so does $\I_s'(\E)$ unless
  extending $\E$ with $\visZ(\eta_1,\eta_2)$ violates
  $\I_c$\footnote{\GK{ToDo: consider other well-formedness conditions
  on trace, such as acyclicity of $\visZ$ and $\soZ$. Are they needed
  (considering that the machine never violates them)? Encode the
  specifications in Z3 and make sure they are consistent.}}:
  \begin{smathpar}
  \begin{array}{l}
  \forall\E,\eta_1,\eta_2.~ \I_c(\E) \Rightarrow (\I_s(\E)
    \Rightarrow \underE{\eta_1 \visar \eta_2}) \Rightarrow \\
    \hspace*{0.5in}(\I_s'(\E) \Rightarrow \underE{\eta_1 \visar
    \eta_2} \disj \neg\I_c(\E\,\cup\,(\emptyset,\{(\eta_1,\eta_2)\})))
  \end{array}
  \end{smathpar}
\end{itemize}
\end{definition}
Applying this principle, we can weaken {\sc sc} to obtain the
following store trace invariant ($\I_s$) for an {\sc sc} store whose
isolation constraints are captured by $\I_c$:
\begin{smathpar}
\begin{array}{lcl}
\I_s(\E) & = & \forall \eta_1,\eta_2.\, \{\eta_1,\eta_2\},
    \subseteq \E.\A \conj \id(\eta_1) <
    \id(\eta_2) \\
    & & \hspace*{0.5in} \Rightarrow 
      \underE{\eta_1 \visar \eta_2} \disj \neg\I_c(\E
    \cup (\emptyset,\{(\eta_1,\eta_2)\}))\\
\end{array}
\end{smathpar}
$\I_s$ requires a trace $\E$ to satisfy the visibility constraints of
{\sc sc} except in cases where they are in conflict with $\I_c$.
Instantiating the parameter $\I$  with $\I_s \wedge \I_c$ in
Fig.~\ref{fig:txnimp} results in an operational semantics that
describes many kinds of relational database-like stores.

For example, the semantics of a causally consistent data
store~\cite{gotsmanpopl16,LBC16} can also be obtained by defining
$\I_s$ to be the causal consistency ({\sc cc}) invariant that is
appropriately weakened to resolve conflicts with $\I_c$.\footnote{Details
of this instantiation are relegated to the supplementary in the interest of space.}

%% \begin{remark}
%% Note that the purpose of maximum visibility principle is to
%% rationalize the observed behaviour of databases in practice. In
%% particular, it is not intended to be a guiding principle to engineer
%% data stores.
%% \end{remark}

% \paragraph{A CC store} A causally consistent data store~\cite{gotsmanpopl16,LBC16}
% allows operations to only witness a causally consistent snapshot of the
% global state. The store-specific invariant obtained by weakening the
% {\sc cc} guarantee to account for conflicts with $\I_c$ is shown
% below (the original {\sc cc} does not contain the $\neg\I_c(\dots)$
% disjuncts): 
% \begin{smathpar}
% \begin{array}{lccl}
% \C{CC}(\E) & \;=\; &  & \forall \eta_1,\eta_2.\, 
%       \E \Vdash \eta_1 \soar \eta_2 \Rightarrow  \underE{\eta_1 \visar
%       \eta_2} \\
%     & & & \hspace*{0.6in}\disj \neg\I_c(\E \cup 
%                 (\emptyset,\{(\eta_1,\eta_2)\}))\\
%     &   & \wedge & \forall\eta_1,\eta_2,\eta_3.\,\underE{\eta_1 \visar
%       \eta_2} \conj \underE{\eta_2 \visoar \eta_3} \\
%     &   & &\hspace*{0.3in} \Rightarrow \underE{\eta_1 \visar \eta_3}
%       \disj \neg\I_c(\E \cup (\emptyset,\{(\eta_1,\eta_2)\}))\\
% \end{array}
% \end{smathpar}

% \noindent Instantiating the parameter $\I$  with $\I_s \wedge \I_c$ in
% Fig.~\ref{fig:txnimp} results in an operational semantics that admits
% violation of causal consistency if and only if the violation is
% inevitable to enforce $\I_c$.

\subsection{Weakly Consistent Replication}
\label{sec:replication}

\begin{figure}
\centering
\subcaptionbox {
  Replica model
  \label{fig:ec-theirs}
} [
  0.4\columnwidth
] {
  \includegraphics[scale=0.7]{Figures/ec-theirs}

}
\hspace*{0.1in}
\subcaptionbox {
  Subset model
  \label{fig:ec-ours}
}{
  \includegraphics[scale=0.45]{Figures/ec-ours}
} \caption{In the replica model, operation $\op$ generates effect
$\eta$ at replica $R_1$, which is then merged to $R_2$. If the
\emph{store is {\sc cc}}, then $R_2$'s state at merge event is same or
larger than $R_1$'s state at generation event (the difference is
highlighted). In our subset model, $\op$ witnesses $S_1 \subseteq
\E.\A$ and generates $\eta$, which is immediately added to $\E.\A$. A
later operation may witness $S_2 \subseteq \E.\A$, and if the
\emph{operation is} {\sc cc} and $\eta \in S_2$, then it also
witnesses $S_1$ (i.e., $S_1 \subseteq S_2$). } 
% Moreover, Like $R_2 - R_1$, if all effects in $S_2 - S_1$ are
% concurrent with $\eta$, i.e., $\not\exists\eta'.~\eta' \in S_2 - S_1
% \conj % visZ(\eta,\eta')$, then any precondition $P$ that is valid
% when $\op$ executed is also valid when $\eta$ is witnessed because
% of the stability condition.
\label{fig:ec-theirs-vs-ours}
\end{figure}

Reasoning under weakly consistent replication has received special
attention in recent work~\cite{gotsmanpopl16}. Our operational
semantics and proof system are general enough to admit replication as
a special case. In this section, we explain how the standard artifacts
of weakly-consistent replication manifest in our reasoning framework.

Reasoning about weakly-consistent replication involves reasoning in
terms of \emph{replicas} of data. The primary challenge in this
setting is to ensure that the assumptions made and guarantees enforced
by an operation at one replica carry over to other replicas that merge
their effects, thus preserving the overall integrity of the system.
In~\cite{gotsmanpopl16}, this challenge is partly addressed by
imposing restrictions on how various replica states differ, i.e., by
fixing a system model with a stronger baseline consistency ({\sc cc})
than {\sc ec}. This unfortunately restricts the reasoning approach
from being applied to stores (e.g., \cite{bayou,pldi15}) that provide
guarantees weaker than causal consistency, such as causal visibility
or read-my-writes~\cite{zoo}. Causal consistency is not baseline
consistency in these stores because it is not \emph{highly
available}~\cite{bailishat}.

Notably, our view of replication however does not explicitly involve
replicas.  Fig.~\ref{fig:ec-theirs-vs-ours} contrasts our model of
weakly-consistent replication with a conventional replica-based
model.  Under our model, the notion of a replica is subsumed by the
concept of visibility; a replica is defined by the subset ($S$) of
global state ($\E.\A$) that an operation witnesses. Constraints over
replica states therefore manifest as constraints over a specific visibility
relation.  For example, instead of requiring the store to be causally
consistent, an operation can request to witness a causally consistent
subset of the state; such demands can be made via the trace invariant
$\I$. For a precondition ($P$) of the operation to be useful, it has
to be an assertion over every causally consistent subset of the global
state.  Since any replica that eventually executes the operation has
to expose one such subset ($S$), the precondition is guaranteed to
hold regardless of the replica. There is however one problem with this
explanation - by considering subsets of just one global state, it
ignores the fact that the global state (hence, the replica states)
change during the execution of the operation. To account for this
change,~\cite{gotsmanpopl16} distinguishes between effect generation
event at one replica $r_1$ and effect merge event at replica $r_2$,
and requires that only \emph{non-conflicting} operations execute
between these two events at $r_2$, and that they preserve certain
invariants. Our framework folds all such machinery into a stability
condition predicated on $\I$ (\S\ref{sec:rely-guarantee}).  Since any
change to the global state during the execution of the operation is an
interference, and $P$ is required to be stable with respect to any
such interference, it follows that $P$ is valid on every replica, thus
ensuring that assumptions made at the generation event are also valid
at the merge event.

\subsection{Example}

We shall now consider the proof of the example in
Fig.~\ref{fig:motiv-eg-1} in greater detail. We assume an {\sc sc}
store, such as a relational database, whose store-specific trace
invaraint ($\I_s$) was shown previously. Both transactions are run at
{\sc si} isolation, hence $\I_c$ is $\lambda\E.~\underE{\C{SI(Wd1)}}
\conj \underE{\C{SI(Wd1)}}$. As usual, $\I$ is $\I_s \conj \I_c$.

\begin{figure}
\centering
\begin{txnimpcode}
$\begin{decoration}
  \hspace*{0.3in}P : \{\neg\committed(\C{Wd1}) \conj \neg\committed(\C{Wd2}) \conj
                          \C{C} = \C{k}\}
\end{decoration}$

 $\begin{decoration}
 P_1:\{ \neg\committed(\C{Wd1}) \conj
        (\neg\committed(\C{Wd2}) \Rightarrow \C{C = k}) \conj\\
        \hspace*{0.3in}{\committed(\C{Wd2})} \Rightarrow 
            \C{Wd2} \wrstoar \C{C} \wedge \C{C = k-a2} \}
 \end{decoration}$
  txn$\langle$'Wd1'$\rangle${
              $\vdots$ 
     $\begin{decoration}
     \phi_4 : \{{\neg\committed(\C{Wd2})} \Rightarrow \C{C = k} \wedge \C{v3 = k-a1} \conj\\
        \hspace*{0.3in}({\committed(\C{Wd2})} \Rightarrow \C{Wd2}
        \wrstoar \C{C}) \conj \neg\committed(\C{Wd1}) \conj \\
        \hspace*{0.32in}{\committed(\C{Wd2})} \wedge
         {\C{Wd2} \visar \C{Wd1}} 
        \Rightarrow \C{C = k-a2} \wedge \\
        \hspace*{1.9in}\C{v3 = k-a2-a1}\}
     \end{decoration}$ 
     C := v3
     $\begin{decoration}
      \phi_5 : \{({\neg\committed(\C{Wd2})} \Rightarrow \C{C = k-a1}) 
          \conj \neg\committed(\C{Wd1}) \conj\\
         \hspace*{0.4in}{\committed(\C{Wd2})} 
            \Rightarrow \C{Wd2} \wrstoar \C{C} \wedge \C{C = k-a1-a2}\}
      \end{decoration}$ 
              $\vdots$ 
  }
 $\begin{decoration}
  Q_1 : \{{(\neg\committed(\C{Wd2})} \Rightarrow \C{C = k-a1})
            \conj \committed(\C{Wd1}) \conj\\
      \hspace*{0.4in}{\committed(\C{Wd2})} 
          \Rightarrow \C{C = k-a1-a2} \}
  \end{decoration}$

$\begin{decoration}
  \hspace*{0.3in}Q : \{\committed(\C{Wd1}) \conj \committed(\C{Wd2}) \conj
                          \C{C} = \C{k-a1-a2}\}
\end{decoration}$
\end{txnimpcode}

\caption{\C{Wd1} transaction decorated with assertions}
\label{fig:wd1-decorated}
\end{figure}

Decorated parts of \C{Wd1} relevant to the discussion are shown in
Fig.~\ref{fig:wd1-decorated}.  \C{Wd2} is similar and not
shown\footnote{Supplementary contains fully decorated program.}.
Temporary name \C{v3} has been introduced to capture the result of the
RHS expression \C{C-a2}.  Assertions implicitly refer to the current
execution ($\E$), just as hoare triples implicitly refer to the
current state.  The context for propositions is also the implicit,
i.e., we write $\psi$ instead of $\underE{\psi}$. The proposition \C{k
$\ge$ a1+a2} remains an invariant, hence elided. The precondition
$P_1$ of \C{Wd1} accounts for the possibility of \C{Wd2} committing
before \C{Wd1}, writing \C{k-a2} to \C{C}. Since neither \C{Wd1} nor
\C{Wd2} are committed at the beginning, the precondition $P$ of the
program is extended with $\neg\committed(\C{Wd1}) \wedge
\neg\committed(\C{Wd2})$, from which $P_1$ follows. The postcondition
$Q_1$ of \C{Wd1} asserts different values for \C{C} depending on
whether or not \C{Wd2} has already committed.  It also asserts that
\C{Wd1} has committed. The postcondition $Q_2$ of \C{Wd2} (not shown)
is similar, but with the roles of \C{Wd1} and \C{Wd2} exchanged. The
postcondition $Q$ follows from $Q_1 \wedge Q_2$, hence valid. It
remains to show that \C{Wd1} satisfies its specification (\C{Wd2} must
also satisfy its spec, but the proof is similar).

First, we focus on the sequential aspect of the proof and show that
the assertions that decorate \C{Wd1} are indeed valid. We consider the
triple associated with the assignment to \C{C} for illustration. The
precondition $\phi_4$ asserts different values for \C{C} and \C{v3}
(temporary alias of RHS expression) depending on whether or not
\C{Wd2} is committed. Inside \C{Wd1}, $\committed(\C{Wd2})$ may mean
that either \C{Wd2} happened before \C{Wd1} (hence can be visible to
all of \C{Wd1}), or that it committed concurrently with \C{Wd1} (hence
cannot be visible to all of \C{Wd2}).  Nonetheless, $\phi_4$ only
considers the case when $\C{Wd2} \visar \C{Wd1}$. As we show below,
this is sufficient to prove the postcondition $\phi_5$, which appears
to make a stronger claim about the value of \C{C}. The proof follows
from the RG rule \rulelabel{RG-Asgn} for assignments, which allows us
to easiliy conclude the following about the execution state after the
assignment:
\begin{smathpar}
\begin{array}{l}
    {\neg\committed(\C{Wd1}) \conj \C{Wd1} \wrstoar \C{C} 
    \conj (\neg\committed(\C{Wd2})} \Rightarrow \C{C = k-a1}) \conj\\
    ({\committed(\C{Wd2})} \Rightarrow \C{Wd2} \wrstoar
    \C{C}) \conj \\
    ({\committed(\C{Wd2})} \wedge {\C{Wd2} \visar \C{Wd1}} 
    \Rightarrow \C{C = k-a2} \wedge \C{C = k-a2-a1})
\end{array}
\end{smathpar}
The rule also lets us assume that the execution satisfies trace
invariant ($\I$), which asserts {\sc si} for both transactions.  Since
both transactions now write to \C{C}, {\sc si} requires either
$\C{Wd1} \visar \C{Wd2}$, or $\C{Wd2} \visar \C{Wd1}$. Since \C{Wd1}
is not yet committed ($\neg\committed(\C{Wd1})$), {\sc si} (which
subsumes {\sc rc}) prohibits the former possibility, allowing us the
deduce $\C{Wd2} \visar \C{Wd1}$. This lets us derive
${\committed(\C{Wd2})}  \Rightarrow \C{C = k-a2} \wedge \C{C =
k-a2-a1}$, allowing us to prove the postcondition $\phi_5$.

The second part of the proof is to show that assertions are stable
despite the interference from the the concurrent thread executing
\C{Wd2}. We formalize the interference as a rely relation that
describes the semantics of \C{Wd2} at just the right level of
abstraction. In particular, we would like to capture two important
facts about \C{Wd2}: 
\begin{itemize}
  \item  If \C{Wd2} commits, then it should have already written to
  \C{C}. This fact is needed to show that \C{Wd2} conflicts with
  \C{Wd1}, so that \C{SI(Wd1)} property can be exploited to deduce
  $\C{Wd2} \visar \C{Wd1}$ from $\committed(\C{Wd2})$ in the
  sequential proof.

  \item The value written to \C{C} is either $\C{k-a1-a2}$ or
  $\C{k-a1}$ depending on whether or not \C{Wd1} has already
  committed. This is clearly necessary to deduce the value of \C{C}
  before and after \C{Wd1} commits.
\end{itemize}
 The resultant rely relation ($R_1$) is shown below (note: we
write $(\E'-\E) \Vdash {\committed(\C{Wd2})}$ to express that \C{Wd2}
is not committed in $\E$, but committed in $\E'$.):

\begin{smathpar}
\begin{array}{lcl}
  R_1 & = & \{ (\E,\E') \;|\; \neg\underE{\committed(\C{Wd1})} \conj
        (\E'-\E) \Vdash {\committed(\C{Wd2})}\\
%       \neg\underE{\committed(\C{Wd2})} \conj \I(\E) \conj \I(\E')\\
%       \conj (\E'-\E) \subseteq \C{Wd2} \conj \\
%   & & \hspace*{0.5in} \E' \Vdash \committed(\C{Wd2}) ~\Rightarrow~ \E'
%       \Vdash \C{Wd2} \wrstoar \C{C} \conj \\ 
%   & & \hspace*{0.5in} \underE{\committed(\C{Wd1})} \conj \E' \Vdash
%   \committed(\C{Wd2}) \Rightarrow \C{C=k-a1-a2} \\
%   & & \hspace*{0.5in} \neg\underE{\committed(\C{Wd1})} \conj
%       (\E'-\E) \Vdash {\committed(\C{Wd2})}\\
    & & \hspace*{0.8in}\Rightarrow \C{C=k-a2} \conj 
        \E' \Vdash \C{Wd2} \wrstoar \C{C} \\
    & & \hspace*{0.37in} \conj \underE{\committed(\C{Wd1})} \conj
        (\E'-\E) \Vdash {\committed(\C{Wd2})}\\
%       \C{COMMIT(Wd2)} \in (\E'.\A-\E.\A) \\
    & & \hspace*{0.6in}\Rightarrow \C{C=k-a1-a2} \conj
        \E' \Vdash \C{Wd2} \wrstoar \C{C} \\
    & & \hspace*{0.37in}\conj \forall(\eta \in \E'.\A - 
        \E.\A).~\txn(\eta) = \C{Wd2}\}\\
\end{array}
\end{smathpar}
For the sake of completeness, we also explicitly represent the fact
that all the effects added by $R_1$ are of \C{Wd2}. This is
needed to prove that an interference from \C{Wd2} does effect the
local state of \C{Wd1}.

The stability for $\phi_4$ w.r.t $R_1$ is trivial to establish: if
$R_1$ does not commit \C{Wd2}, then the value of \C{C} is unaffected,
and if it commits \C{Wd2}, then $\C{Wd2}\not\visar\C{Wd1}$, hence
$\phi_4$ doesn't care. For $\phi_5$, {\sc si} condition should be used
to show that any interference from \C{Wd2} is invalid. The proof
proceeds on the same lines as the Hoare proof discussed above; it is
also discussed in detail in \S\ref{sec:motivation}.

The final aspect of the proof is to show that any interference from
\C{Wd1} is contained in its guarantee relation $G_1$, which is also
the rely relation $R_2$ for \C{Wd2}. Since the implementation of
\C{Wd2} is same as \C{Wd1}, $R_2$ (hence $G_1$) is same as $R_1$ shown
above, but with \C{Wd2} and \C{Wd1} interchanged. Continuing our focus
on the assigment statement, the guarantee proof requires us to show
that if executing the assignment takes the execution from $\E$ to
$\E'$, such that $\E'$ satisfies $\I$, then $(\E,\E') \in G_1$.  The
proof follows from the \rulelabel{RG-Asgn} rule, which establishes
that (a). the assignment statement does not commit \C{Wd2}, and (b).
it adds a write effect $\eta$ such that $\txn(\eta)=\C{Wd1}$. This
finishes the proof for the triple in focus. The proofs for other
triples of \C{Wd1} can also be constructed similarly. The combination
of correctness proofs for \C{Wd1} and \C{Wd2} yeilds a correctness
proof for the program.