### 2016-07-04 case-studies.tex

\section{Case Studies}
\label{sec:case-studies}

In this section, we present two case studies that demonstrate the
applicability of our reasoning approach in varied settings, and its
consequent practical utility.

\paragraph{A video view counter.} Our first application is a
Ruby-on-Rails implementation of a counter to count the number of views
a video has garnered on a video-sharing website like
YouTube\footnote{Source code of the applications, experiments, and the
  tools we developed to analyze Ruby can be found at the following
  anonymized url: \url{...}}. The application stores the video count
in a MySQL database, and supports a read operation (a database read
transaction), and an increment operation (a read-increment-write
sequence wrapped in a transaction). The application is expected to
preserve the \emph{monotonicity invariant}, namely that the view count
should never appear to be decreasing. However, the default
\iso{Repeatable Read} isolation level of MySQL leads to anomalous
executions - with 32 concurrent writers performing single increments,
and one reader issuing a constant stream of reads, we observed 12
violations of monotonicity in 10 rounds. We subsequently instantiated
our reasoning framework for an {\sc sc} store with ANSI SQL isolation
levels, and spent 5 man hours determining appropriate isolation levels
needed to enforce the invariant.\footnote{The representative \txnimp
  programs we used to formally reason about scenarios described in
  this section can be found in the supplementary.} In particular, we
were able to formally prove that to preserve the invariant, writers
need to execute using \iso{Serializable} isolation, while readers can
execute using \iso{Read Uncommitted}. Repeating the experiments with
this configuration led to no violations, as expected. While executing
all transactions at {\sc ser} also prevents violations, it led to an
additional 113\% increase in the latency of writes on average, using a
configuration with 32 concurrent writers, and 5 rounds. We repeated
the experiments with the same configuration on Postgres, and found no
violations. However, Postgres' isolation levels~\cite{postgresiso} are
stronger than ANSI SQL's in non-trivial ways. We therefore formalized
Postgres' guarantees and repeated relevant parts of the proof to
discover that \iso{Repeatable Read} is in fact sufficient for writers
- an observation confirmed through experiments. As usual, using {\sc
  ser} for all transactions also suffices, but on Postgres results in
an additional 47\% increase in write latency.

% To understand whether a straightforward implementation of the view
% counter preserves the invariant, we considered a \txnimp program
% that is representative of the counter application. In particular, we
% considered four concurrent transactions that operate on a shared
% counter: a reader transaction that reads the counter twice obtaining
% the values $k_1$ and $k_2$, respectively, and three writer
% transactions, each incrementing the counter once. If the reader
% witnesses a monotonically increasing view count, then $k_2\ge k_1$.
% We then instantiated our reasoning framework for an {\sc sc} store
% with {\sc rc} as the default isolation level (Postgres is one such
% store), with the aim of verifying the program under this
% instantiation.  Our attempts were however unsuccessful owing to our
% inability to prove that a writer transaction always commits a value
% no smaller than the current value of the counter. This points to the
% existence of anamolous executions where  To understand the
% prevalence of such anomalies, we implemented a view counter
% application in Ruby-on-Rails using standard templates and
% techniques.   
To understand the effect of changing store consistency levels, we
re-implemented the counter as an op-based \emph{replicated data type}
(RDT)~\cite{crdt,burckhardt14} on top of Cassandra
\emph{bolted-on}~\cite{bailisbolton} with {\sc cc} and {\sc sc}
consistency levels (both on-demand; {\sc ec} is default), and ANSI SQL
isolation levels. Keeping the consistency level at {\sc ec}, we then
repeated our experiments with committed reads and serializable
increments - a combination which was proved to preserve monotonicity
on an {\sc sc} store, but nonetheless observed monotonicity
violations. The reason for such violations became evident when
we instantiated our proof framework for an {\sc ec} store.
Under this instantiation, it is impossible to derive the
\C{MonotonicVis} (\S\ref{sec:ansi-isolation}) property needed to show
that successive reads in a session witness monotonically increasing
state. However, assuming {\sc cc} store consistency for reads allowed
us to complete the proof, and recover the monotonicity invariant. In
fact, we were able to complete the proof even after relaxing the
serializability requirement for increments, just by changing the
definition of the $\interp{\E}(X)$ (\S\ref{sec:opsem}) to return the
value of the largest write-to-$X$ instead of the latest write-to-$X$.
In other words, if the reads chose the largest valued write effect,
instead of the latest write effect they witness, then counter
increments need not be serializable for reads to witness a
monotonically increasing count.\footnote{The trade-off is that the view
count is now no longer accurate, although it grows monotonically.}
We implemented this version of the counter as an op-based RDT, which
uses {\sc cc} reads and {\sc ec} increments, both un-isolated (i.e.,
\emph{Read Uncommitted}), and obtained a 52\% improvement in overall
latency compared to the version that uses serializable increments.
\vspace*{-3pt}
\paragraph{Microblog.} Our second set of experiments focused on a Twitter-like microblogging
site, also written using Ruby-on-Rails~\cite{railsbook}.  An important
invariant that must be preserved is referential
integrity between microposts and authors. The application enforces
this invariant entirely in user space by relying on database-backed
transactions (this behavior is typical of Rails
applications~\cite{bailisferal}).  For instance, deletion of a user is
carried out in a transaction that also includes the deletion of
microposts authored by the user.  However, it is possible for concurrent observers 
to nonetheless witness invariant violations. To measure
the likelihood of such anomalies, we first populated a Postgres
database with 1000 user accounts, each user with 50 microposts and 20
followers, chosen uniformly at random. We then constructed experiments
with 64 concurrent clients performing deletions and timeline reads of
randomly chosen users in a 1:7 ratio, and witnessed 250 violations of
referential integrity. To determine appropriate isolation levels, we
then instantiated our framework for an {\sc sc} store with various
ANSI SQL isolation levels for user deletions and timeline reads, and
obtained a proof that {\sc rc} is sufficient for deletions, whereas
reads require {\sc rr}. The proof is immediately applicable for MySQL,
which implements ANSI SQL isolation. Repeating our experiments
for MySQL with this configuration resulted in no violations (as
expected), while yielding a 38\% reduction in latency than an {\sc
  ser} configuration (96 concurrent clients; 5 rounds).  Since
Postgres isolation levels are stronger than the corresponding ANSI
levels, \rulelabel{RG-Conseq} rule (Fig.~\ref{fig:rg-rules}) lets us
immediately deduce that the same choice of isolation levels also work
on Postgres.

Clearly, there exists a significant diversity among the weak
consistency and isolation variants implemented on various real-world
data stores. More such variants have been developed in recent years to
meet the need for high availability (\emph{e.g.}, \emph{Parallel
  Snapshot Isolation}~\cite{psi} (2011) or \emph{Non-Monotonic
  Snapshot Isolation}~\cite{nmsi} (2013)).  As more commercial
weakly-consistent stores adopt various kinds of transactional
semantics, like Cassandra's recently introduced \emph{lightweight
  transactions}, it is reasonable to expect new isolation
definitions to continue to be proposed.  While it may be possible to
carefully engineer a reasoning framework for each combination of
consistency and isolation, such a strategy would lead to multiple
semantic definitions and proof systems with no obvious way to compare
and relate them. As illustrated here, having a parameterized semantics
and a proof framework built on top of it allows us to support these
variants as distinct, yet comparable, instantiations.

%  GK: what does the following sentence mean?
%  As a consequence, we avoid the need for semantics re-engineering
%  every time an application is retargeted to a new platform.