### 2016-07-06 case-studies.tex
Github link added.

\section{Case Studies}
\label{sec:case-studies}

In this section, we present two case studies that demonstrate the
applicability of our framework in varied settings, and its consequent
practical utility. Both our case studies target Ruby-on-Rails
applications. To analyze these, and other similar open-source
applications, we have developed a symbolic execution engine that
compiles a Rails program to an abstract program in an extended version
of \txnimp, along with some proof automation support.\footnote{Details
   can be found in the supplementary material.  Source code for our
  implementation, along with applications and experiments, are
  available at the following anonymized url:
  \url{https://github.com/popl17/popl17-code}} The focus of
our formal verification efforts are such translated \txnimp programs.

\paragraph{A video view counter.} Our first application is a Rails
implementation of a counter to count the number of views a video has
garnered on a video-sharing website like YouTube. The application
stores the video count in a MySQL database, and supports a read
operation (a database read transaction), and an increment operation (a
read-increment-write sequence wrapped in a transaction).  The
application is expected to preserve a \emph{monotonicity invariant},
namely that the view count should never appear to be decreasing.
However, the default \iso{Repeatable Read} isolation level of MySQL
leads to anomalous executions - with 32 concurrent writers performing
single increments, and one reader issuing a constant stream of reads,
we observed 12 violations of monotonicity in 10 rounds. We
subsequently instantiated our reasoning framework for an {\sc sc}
store with ANSI SQL isolation levels, and spent 5 man hours
determining appropriate isolation levels needed to enforce the
invariant. In particular, we were able to formally prove that to
preserve the invariant, writers need to execute using
\iso{Serializable} isolation, while readers can execute using
\iso{Read Uncommitted}. Repeating the experiments with this
configuration led to no violations, as expected. While executing all
transactions at {\sc ser} also prevents violations, it led to an
additional 113\% increase in the latency of writes averaged over 5
rounds with 32 concurrent writers.  We repeated the experiments with
the same configuration on Postgres, and found no violations. However,
Postgres' isolation levels~\cite{postgresiso} are stronger than ANSI
SQL's in non-trivial ways. We therefore formalized Postgres'
guarantees and repeated relevant parts of the proof to discover that
\iso{Repeatable Read} is in fact sufficient for writers - an
observation further validated through experiments. As usual, using
{\sc ser} for all transactions also suffices, but on Postgres results
in an additional 47\% increase in write latency.

% To understand whether a straightforward implementation of the view
% counter preserves the invariant, we considered a \txnimp program
% that is representative of the counter application. In particular, we
% considered four concurrent transactions that operate on a shared
% counter: a reader transaction that reads the counter twice obtaining
% the values $k_1$ and $k_2$, respectively, and three writer
% transactions, each incrementing the counter once. If the reader
% witnesses a monotonically increasing view count, then $k_2\ge k_1$.
% We then instantiated our reasoning framework for an {\sc sc} store
% with {\sc rc} as the default isolation level (Postgres is one such
% store), with the aim of verifying the program under this
% instantiation.  Our attempts were however unsuccessful owing to our
% inability to prove that a writer transaction always commits a value
% no smaller than the current value of the counter. This points to the
% existence of anamolous executions where  To understand the
% prevalence of such anomalies, we implemented a view counter
% application in Ruby-on-Rails using standard templates and
% techniques.   
To understand the effect of changing store consistency levels, we
re-implemented the counter as an op-based \emph{replicated data type}
(RDT)~\cite{crdt,burckhardt14} on top of Cassandra
\emph{bolted-on}~\cite{bailisbolton} with {\sc cc} and {\sc sc}
consistency levels (both on-demand; {\sc ec} is default), and ANSI SQL
isolation levels. Keeping the consistency level at {\sc ec}, we then
repeated our experiments with committed reads and serializable
increments - a combination which was proved to preserve monotonicity
on an {\sc sc} store, but nonetheless observed monotonicity
violations. The reason for such violations became evident when
we instantiated our proof framework for an {\sc ec} store.
Under this instantiation, it is impossible to derive the
\C{MonotonicVis} (\S\ref{sec:ansi-isolation}) property needed to show
that successive reads in a session witness monotonically increasing
state. However, assuming {\sc cc} store consistency for reads allowed
us to complete the proof, and recover the monotonicity invariant. In
fact, we were able to complete the proof even after relaxing the
serializability requirement for increments, just by changing the
definition of the $\interp{S}(X)$ (\S\ref{sec:opsem}) to return
the value of the largest write-to-$X$ instead of the latest
write-to-$X$ in $S$.  In other words, if the reads chose the largest
valued write effect, instead of the latest write effect they witness,
then counter increments need not be serializable for reads to witness
a monotonically increasing count.\footnote{The trade-off is that the
view count is now no longer accurate, although it grows
monotonically.} We implemented this version of the counter as an
op-based RDT, which uses {\sc cc} reads and {\sc ec} increments, both
un-isolated (i.e., \emph{Read Uncommitted}), and obtained a 52\%
improvement in overall latency compared to the version that uses
serializable increments.
\vspace*{-3pt}

\paragraph{Microblog.} Our second set of experiments focused on a
Twitter-like microblogging application taken from~\cite{railsbook}.
An important invariant that must be preserved is referential integrity
between microposts (tweets) and authors. The application enforces this
invariant entirely in user space by relying on database-backed
transactions (this behaviour is typical of Rails
applications~\cite{bailisferal}).  For instance, the deletion of a
user is carried out in a transaction that also includes the deletion
of microposts authored by the user.  However, it is possible for
concurrent observers to nonetheless witness invariant violations. To
quantify the likelihood of such anomalies, we first populated a
Postgres database with 1000 user accounts, in which each user is
associated with 50 microposts and 20 followers, chosen uniformly at
random. We then constructed experiments with 64 concurrent clients
performing deletions and timeline reads of randomly chosen users in a
1:7 ratio, and witnessed 250 violations of referential integrity (the
isolation level was left at the default {\sc rc} level for these
experiments). To determine appropriate (correct) isolation levels, we
then instantiated our framework for an {\sc sc} store with various
ANSI SQL isolation levels for user deletions and timeline reads, and
obtained a proof that {\sc rc} is sufficient for deletions, whereas
reads require {\sc rr}. The proof is immediately applicable for MySQL,
which implements ANSI SQL isolation levels. Repeating our experiments
for MySQL with this configuration resulted in no violations (as
expected), while yielding a 38\% reduction in latency than an {\sc
  ser} configuration.  Since Postgres' isolation levels are stronger
than the corresponding ANSI levels, \rulelabel{RG-Conseq} rule
(Fig.~\ref{fig:rg-rules}) lets us immediately deduce that the same
choice of isolation levels also work on Postgres.

Clearly, there exists significant diversity among the weak consistency
and isolation variants implemented on various real-world data
stores. More such variants have been developed in recent years to meet
the need for high availability (\emph{e.g.}, \emph{Parallel Snapshot
  Isolation}~\cite{psi} (2011) or \emph{Non-Monotonic Snapshot
  Isolation}~\cite{nmsi} (2013)).  As more commercial
weakly-consistent stores adopt various kinds of transactional
semantics, like Cassandra's recently introduced \emph{lightweight
  transactions}, it is reasonable to expect new isolation definitions
to continue to be proposed.  While it may be possible to carefully
engineer a reasoning framework for each combination of consistency and
isolation, such a strategy would lead to multiple semantic definitions
and proof systems with no obvious way to compare and relate them. As
illustrated here, having a parameterized semantics and a proof
framework built on top of it allows us to support these variants as
distinct, yet comparable, instantiations.