### 2017-07-02 abstract.tex

\begin{abstract}

  Serializability is a well-understood correctness criterion that
  simplifies reasoning about the behaviour of concurrent transactions
  by ensuring they are \emph{isolated} from each other while they
  execute.  However, enforcing serializable isolation comes at a steep
  cost in performance because it necessarily restricts opportunities
  to exploit concurrency even when such opportunities would not
  violate application-specific invariants. As a result, database
  systems in practice support, and often encourage, developers to
  implement transactions using weaker alternatives. These alternatives
  break the strong isolation guarantees offered by serializable
  transactions to permit greater concurrency. Unfortunately, the
  semantics of weak isolation is poorly understood, and usually
  explained only informally in terms of low-level implementation
  artifacts. Consequently, verifying high-level correctness properties
  in such environments remains a challenging problem.

  To address this issue, we present a novel program logic that enables
  compositional reasoning about the behavior of concurrently executing
  weakly-isolated transactions.  Recognizing that the proof burden
  necessary to use this logic may dissuade application developers, we
  also describe a fully automated verification tool based on this
  foundation that determines the safety of high-level consistency
  assertions associated with such transactions.  The key to effective
  inference is the observation that weakly-isolated transactions can
  be viewed as functional (monadic) computations over database state,
  allowing us to treat their operations as state transformers over the
  database.  This interpretation enables automated verification using
  off-the-shelf SMT solvers.

  Notably, our development is parametric over a transaction's specific
  isolation semantics, allowing it to be applicable over a range of
  concurrency control mechanisms.  Case studies and experiments on
  real-world applications demonstrate the utility of our approach, and
  provide strong evidence that automated verification of
  weakly-isolated transactions can be placed on the same formal
  footing as their strongly-isolated serializable counterparts.

\end{abstract}