### 2017-07-02 opsem.tex
Edits.

\section{\txnimp: Syntax and Semantics}
\label{sec:opsem}

\label{sec:syntax}

\input{txnimp}

Fig.~\ref{fig:txnimp} shows the syntax and small-step semantics of
\txnimp, a core language that we use to formalize our intuitions about
reasoning under weak isolation. Variables ($x$), Integer and boolean
constants ($k$), records ($r$) of named constants, sets ($s$) of such
records, arithmetic and boolean expressions ($e_1 \odot e_2$), and
record expressions ($\{\bar{f}=\bar{e}\}$) constitute the syntactic
class of expressions ($e$). Commands ($c$) include $\cskip$,
conditional statements, \C{LET} constructs to bind names, \C{FOREACH}
loops, SQL statements, their sequential composition ($c_1;c_2$), and
transactions ($\ctxn{i}{\I}{c}$) and their parallel composition
($c_1\,||\,c_2$). The $\I$ in \C{TXN} block syntax is the
transaction's isolation specification, which will be explained later.
Certain terms that only appear at run-time are also present in $c$.
These include a \C{TXN} block tagged with sets ($\stl$ and $\stg$) of
records that assume special meaning in operational semantics, and a
\C{FOREACH} loop that keeps track of the set ($s_1$) of items already
iterated, and the set ($s_2$) of items yet to be iterated. Note that
even the surface-level \C{FOREACH} shown here is a little different
from the one used in previous sections; its higher-order argument has
two arguments, $y$ and $z$, which are invoked (during the reduction)
with the set of already-iterated items, and the current item,
respectively. This form of \C{FOREACH} lends itself to inductive
reasoning, facilitating inductive proofs (Sec.~\ref{sec:reasoning}).

% also We let $T_i$ for $i \in \mathbb{N}$
% range over transaction identifiers. When it is evident we are
% referring to a transaction, we use the number $i$ instead of $T_i$ for
% identification (\eg in $\C{txn}\langle i \rangle$). Like variables,
% transaction identifiers are globally accessible. For notational
% convenience, we let $t$ range over both expressions and commands.

We define a small-step operational semantics for this language in
terms of an abstract machine that executes a command, and updates
either a transaction-local database ($\stl$), or the global database
($\stg$). Database ($\stg$) is a modeled as a set of records of a
pre-defined type, i.e., they all belong to a single table.
Generalization to multiple tables is straightforward. Records in
$\stg$ are uniquely identifiable through their $\idf$ field, which is
auto-generated and does not belong to the surface language, i.e.,
$\C{id}\notin f$. For a set $S$ of records, we define $\dom(S)$ as the
set of unique ids of all the records in $S$. Thus $|\dom(\stg)| =
|\stg|$. During its execution, a transaction may write to multiple
records in $\stg$. Atomicity dictates that such writes shouldn't be
available in $\stg$ until the transaction commits. We therefore
associate each transaction with a local database ($\stl$) that stores
\emph{only} the uncommitted records\footnote{While SQL's \C{UPDATE}
admits writes at the granularity of record fields, databases, in
reality, enforce record-level locking, allowing us to think of
``uncommitted writes'' as ``uncommitted records''. }. Uncommitted
records include deleted records, which are identified by a hidden
$\delf$ field set to \C{true}. When the transaction commits, its local
database is atomically \emph{flushed} to the global database,
committing the uncommitted records. The flush operation ($\gg$) is defined as
following:
\begin{smathpar}
\begin{array}{c}
\forall r.~ r \in (\stl\gg\stg) ~\Leftrightarrow~ 
  (r.\idf \notin \dom(\stl) \conj r \in \stg)
\disj (r \in \stl \conj \neg r.\delf) 
\end{array}
\end{smathpar}
Let $\stg' = \stl\gg\stg$. A record $r$ belongs to $\stg'$ iff it
belongs to $\stg$ and not been updated in $\stl$, i.e., $r.\idf \notin
\dom(\stl)$, or it belongs to $\stl$, i.e., it is either a new record,
or an updated version of an old record, but the updatation is not a
deletion ($\neg r.\delf$). Thus, flush defines the result of
atomically applying transaction's local writes to the global database.
Besides the commit, flush also helps a transaction read its own
writes. Intuitively, the result of a read operation inside a
transaction must be computed on the database resulting from flushing
the current local state ($\stl$) to the global state ($\stg$). The
abstract machine of Fig.~\ref{fig:txnimp}, however, doesn't let a
transaction read its own writes. This reduces verbosity and simplifies
semantics, without losing any generality; substituting $\stl\gg\stg$
for $\stg$ at select places in reduction rules recovers the real

Small-step semantics is stratified into a transaction-local reduction
relation, and a top-level reduction relation. Transaction-local
relation ($\stg \vdash (c,\stl) \stepsto (c',\stl')$) defines a
small-step reduction for a command inside a transaction, when the
database state is $\stg$; the command $c$ reduces to $c'$, while
updating the transaction-local database $\stl$ to $\stl'$. The
definition assumes a meta function $\eval$ that evaluates expressions
with no free variables to values. The reduction relation for SQL
statements is defined straightforwardly.  \C{INSERT} adds a new record
to $\stl$ after adding a unique identifier (more discussion on
uniqueness later). \C{DELETE} finds the records in $\stg$ that match
the search criteria defined by its boolean function argument, and adds
the records to $\stl$ after marking them for deletion. \C{SELECT}
bounds the name introduced by \C{LET} to the set of records from
$\stg$ that match the search criteria, and then executes the bound
command $c$. \C{UPDATE} uses its first function argument to compute
the updated version of the records that match the search criteria
defined by its second function argument. Updated records are added to
$\stl$.

The reduction of \C{FOREACH} starts by first converting it to its
run-time form that also keeps track of the iterated items ($s_1$),
besides the yet-to-be-iterated items ($s_2$). Initially, $s_1$ is
empty. As the elements are iterated, they are removed from $s_2$ and
added to $s_1$ . Iteration involves invoking the higher-order function
with $s_1$ and the current element $x$ (note: $\uplus$ in $\{x\}
\uplus s_2$ denotes a disjoint union). The reduction ends when the
$s_2$ becomes empty. The reduction rules for conditionals, \C{LET}
binders, and sequences are straightforward, hence ommitted.

The top-level reduction relation defines small-step semantics of
transactions, and their parallel composition. A transaction always
comes tagged with an isolation specification $\I$ that dictates the
timing and nature of iterferences that the transaction can witness.
Formally, $\I$ is a function that maps a boolean ($\mathbb{B}$) and a
triplet of database states (three sets of records:
$\Pow{r}\times\Pow{r}\times\Pow{r}$) to a logical proposition
($\Prop$). The abstract machine calls $\I$ while reducing
$\ctxn{i}{\I}{c}$. The boolean argument identifies if the transaction
is still executing (\C{true}), or if it has finished execution and
ready to commit (\C{false}).  Among the triplet of sets (the second
argument), first is the (current) transaction-local database state
($\stl$), second is the state ($\stg$) of the global database when the
transaction last took a step, and third is the current state ($\stg'$)
of the global database.  Intuitively, $\stg'\neq\stg$ indicates an
interference from a concurrent transaction, and the $\I$ decides if
this interference is allowed or not, considering the local database
state ($\stl$). For instance, as described in
(Sec.~\ref{sec:motivation}), an RR transaction on Postgres defines
$\I$ as following:
\begin{smathpar}
\begin{array}{lcl}
\I\,\C{true}\,(\stl,\stg,\stg') & = & \stg' = \stg\\
\I\,\C{false}\,(\stl,\stg,\stg') & = & \forall(r\in\stl)(r'\in\stg).~
  r'.\idf = r.\idf \Rightarrow r'\in\stg'.
\end{array}
\end{smathpar}

The reduction of a $\ctxn{i}{\I}{c}$ begins by first converting it to
its run-time form $\ctxn{i}{\I,\stl,\stg}{c}$, where $\stl =
\emptyset$, and $\stg$ the current (global) database. The top-level
reduction rules in Fig.~\ref{fig:txnimp} assume this form. Rule
\rulelabel{E-Txn} reduces $\ctxn{i}{\I,\stl,\stg}{c}$ under a database
state ($\stg'$), only if the isolation specification ($\I$) allows the
interference between $\stg$ and $\stg'$ during the execution. Rule
\rulelabel{E-Commit} commits the transaction
$\ctxn{i}{\I,\stl,\stg}{c}$ by flushing its uncommitted records to the
database. This is done only if the interference between $\stg$ and
$\stg'$ is allowed during the commit by the isolation specification.

\input{isolation}