### 2017-07-04 isolation.tex
Edits.

\subsection{Isolation Specifications}
\label{sec:isolation}

A distinctive characteristic of our development is that it is
parameterized on the isolation specification $\I$. $\I$ can be
instantiated with the declarative characterization of an isolation
guarantee or a concurrency control mechanism, regardless of its actual
implementation. This allows us to model a range of isolation
properties that are relevant to the theory and practice of database
systems. We present some below:

% Databases implement isolation levels through a combination of various
% concurrency control mechanisms, which are absent from
% Fig.~\ref{fig:txnimp}. In this section, we first axiomatize the
% semantics of various concurrency control constructs, and then define
% the precise semantics of the isolation levels as they are implemented
% on two real-world databases -  PostgreSQL and MySQL.
% Any additional concurrency control provided by the database
% implementation, beyond that which is already incuded in the
% isolation specification, also needs to be axiomatized in $\I$. We
% first describe such axiomatizations, and then present the
% specifications of the standard isolation levels as implemented by
% popular off-the-shelf databases.

\textbf{Unique Ids}. As the \C{new\_order} example
(Sec.~\ref{sec:motivation}) demonstrates, enforcing global uniqueness
of ordered identifiers requires stronger isolation. Alternatively,
globally unique sequence numbers, regardless of the isolation level,
can be from a requested from a relational databases via SQL's
\C{UNIQUE} and \C{AUTO\_INCREMENT} keywords. Our development crucially
relies on the uniqueness of record identifiers\footnote{This has
parallels in real-world. For e.g., MySQL's InnoDB engine automatically
adds a 6-byte unique identifier if none exists for a record.}, which
are generated locally by the \rulelabel{E-Insert} rule. Global
uniqueness of locally generated identifiers can be axiomatized as an
isolation property as following:
\begin{smathpar}
\begin{array}{lcl}
  \I_{id}(\stl,\stg,\stg') & = & \forall(r\in\stl).~
      r.\idf\notin \dom(\stg) \Rightarrow r.\idf\notin \dom(\stg').
\end{array}
\end{smathpar}
$\I_{id}$ ensures that a globally unique id generated during a
transaction's execution remains globally unique until it commits,
prohibiting interference from a concurrent transaction that adds the
same id. The axiom thus simulates a global counter protected by an
exclusive lock.

\textbf{Write-Write Conflicts}. Databases often employ a combination of
concurrency control methods, both optimistic (e.g., speculation and
rollback) and pessimistic (e.g., exclusive locking), to eliminate
write-write conflicts among concurrent transactions. The absence of
write-write (ww) conflicts, regardless of how it is achieved, can be
axiomatized as following:
\begin{smathpar}
\begin{array}{lcl}
  \I_{ww}(\stl,\stg,\stg') & = & \forall(r'\in\stl)(r \in \stg).~
      r.\idf = r'.\idf  \Rightarrow r\in\stg'.
\end{array}
\end{smathpar}
That is, given a record $r'\in\stl$, if there exists an $r\in\stg$
with the same id (i.e., $r'$ is an updated version of $r$), then $r$
must be present unmodified in $\stg'$. This prevents a concurrent
transaction from changing $r$, thus simulating the behavior of an
exclusive lock or a speculative execution that succeeded (Note: a
transaction writing to $r$ always changes $r$ because its $\txnf$
field is updated). 
% There is however a caveat: if we assume extensional
% equality over records, a write by a concurrent transaction that
% doesn't change $r$'s contents is still allowed. While this in itself
% is not a problem, the concurrent transaction may modify other records,
% which then become visible in the current transaction. A write lock
% prevents this behavior, whereas our axiomatization ($\I_{ww}$) allows
% it. An easy fix for this is to add version timestamps to records that
% effectively intensionalizes equality. Nonetheless, imprecise
% axiomatization of write locks hasn't been a problem in practice.

\textbf{Snapshots} Almost all major relational databases implement
isolation levels that execute transactions against a static snapshot
of the database.  The semantics of a snapshot can be axiomatized as
following:
\begin{smathpar}
\begin{array}{lcl}
  \I_{ss}(\stl,\stg,\stg') & = & \stg' = \stg.
\end{array}
\end{smathpar}
% The snapshot semantics however may not be applicable throughout the
% lifetime of the transaction, like in the case of Repeatable Read (RR).

\textbf{Read-Only Transactions}. Certain databases implement special
previleges for read-only transactions. Read-only behavior can be
enforced on a transaction by including the following proposition in
its $\I$:
\begin{smathpar}
\begin{array}{lcl}
  \I_{ro}(\stl,\stg,\stg') & = & \stl = \emptyset\\
\end{array}
\end{smathpar}
If a transaction declared as read-only performs a write, then its
$\stl\neq \emptyset$, and the transaction never commits.

We now define various isolation guarantees that are either found in
the database and distributed systems literature, or implemented by a
commercial vendor.

\textbf{Read Committed (RC) and Monotonic Atomic View (MAV)} RC
isolation allows a transaction to witness only the writes of the
committed transactions. In particular, \emph{dirty writes} of
uncommitted transactions are not made visible. Monotonic Atomic View
(MAV)~\cite{hat,quelea} is an extension to RC that guarantees the
continuous visibility of a committed transaction's writes if they once
become visible in the current transaction. That is, an MAV transaction
does not witness \emph{disappearing writes}, which can happen on a
weakly consistent machine. Due to the SC nature of our abstract
machine (there is always a single global database state $\stg$; not a
vector of states indexed by vector clocks), and our choice to never
violate atomicity of a transaction's writes, both RC and MAV are
already guaranteed by our semantics. Thus $\I_e$ and $\I_c$ are both
$true$ for RC and MAV.

\textbf{Repeatable Read (RR)} By definition, a read in a Repeatable
Read transaction is required to return the same value regardless of
how many ever times it is repeated. RR is often implemented (for e.g.,
in ~\cite{mysql,quelea,hat}) by executing the transaction against a
(conceptual) snapshot of the database, but committing its writes to
the actual database. This implementation of RR can be axiomatized as
$\I_e = \I_{ss}$ and $\I_{c}=true$. However, this specification of RR
is stronger than the ANSI SQL specification, which requires no more
than the invariance of already read records. In particular, ANSI SQL
RR allows \emph{phantom reads} phenomenon, where a \C{SELECT} query
might return newer records that were not previously present. This
specification is implemented, for e.g., in~\cite{mssql}, using
record-level exclusive read locks, that prevent a record from being
modified while it is read by an uncommitted transaction. ANSI SQL RR
can be axiomatized in our framework, but it requires a minor extension
to our operational semantics to track a transaction's reads. In
particular, the records returned by \C{SELECT} should be added to the
local database $\stl$, but without changing their transaction
idenfitiers ($\txnf$ fields), and flush ($\gg$) should only flush the
records that bear the current transaction's identifier. With this
extension, ANSI SQL RR can be axiomatized thus ($i$ is the id of the
current transaction):
\begin{smathpar}
\begin{array}{lcl}
  \I_e(\stl,\stg,\stg') & \Leftrightarrow & \forall(r'\in\stl)(r \in \stg).~
      r'.\idf \neq i \wedge r.\idf = r'.\idf  \Rightarrow r\in\stg'.\\
  \I_c(\stl,\stg,\stg') & \Leftrightarrow & true\\
\end{array}
\end{smathpar}

\textbf{Snapshot Isolation (SI)} The concept of executing a
transaction against a consistent snapshot of the database was first
proposed as Snapshot Isolation in~\cite{berenson}. SI doesn't admit
write-write conflicts, and the original proposal, which is implemented
in MS SQL Server~\cite{sqlserver}, required the
database to roll-back an SI transaction if conflicts are detected
during the commit. This behavior can be axiomatized as following:
\begin{smathpar}
\begin{array}{lcl}
\I_e\,\,(\stl,\stg,\stg') & = & \stg' = \stg\\
\I_c\,\,(\stl,\stg,\stg') & = & \forall(r\in\stl)(r'\in\stg).~ r'.\idf = r.\idf \Rightarrow r'\in\stg'.
\end{array}
\end{smathpar}
Observe that the same axiomatization applies to Postgres's RR,
although its implementation (described in Sec.~\ref{sec:motivation})
differs considerably from the original proposal. Thus, reasoning done
for an SI transaction on MS SQL server carries over Postgres's RR and
vice-versa, demonstrating the benefits of reasoning axiomatically
about isolation properties.

\textbf{Serializability (SER)} Axiomatization of serializability is
straightforward:
\begin{smathpar}
\begin{array}{lcl}
\I_e\,\,(\stl,\stg,\stg') & = & \stg' = \stg\\
\I_c\,\,(\stl,\stg,\stg') & = & \stg' = \stg.
\end{array}
\end{smathpar}

%\usepackage[table,xcdraw]{xcolor}
\begin{table}[]
\centering
\begin{tabular}{l|l|l|l|l|l|l|l|l|l|l|l|l|}
\cline{2-13}
                                & \multicolumn{6}{c|}{{\color[HTML]{333333} PostgreSQL}}                         & \multicolumn{6}{c|}{MySql}                                                   \\ \cline{2-13} 
                                & \multicolumn{2}{c|}{RC} & \multicolumn{2}{c|}{RR} & \multicolumn{2}{c|}{SER} & \multicolumn{2}{c|}{RC} & \multicolumn{2}{c|}{RR} & \multicolumn{2}{c|}{SER} \\ \cline{2-13} 
                                & T          & F          & T          & F          & T           & F          & T          & F          & T          & F          & T           & F          \\ \hline
\multicolumn{1}{|l|}{$\I_{ss}$} &            &            & \checkmark &            & \checkmark  & \checkmark &            &            & \checkmark &            & \checkmark  & \checkmark \\ \hline
\multicolumn{1}{|l|}{$\I_{rb}$} &            &            &            & \checkmark &             &            &            &            &            &            &             &            \\ \hline
\multicolumn{1}{|l|}{$\I_{ro}$} &            &            &            &            &             &            &            &            & \checkmark & \checkmark &             &            \\ \hline
\multicolumn{1}{|l|}{$\I_{id}$} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark \\ \hline
\multicolumn{1}{|l|}{$\I_{ww}$} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark  & \checkmark \\ \hline
\end{tabular}

\caption{The semantics of isolation levels on PostgreSQL and MySQL}
\label{fig:iso-table}
\end{table}

With help of the above axiomatizations, we can now define the precise
semantics of various isolation levels on PostgreSQL and MySQL. The table
in Table~\ref{fig:iso-table} captures the summary.