### 2017-07-04 abstract.tex

\begin{abstract}

  Serializability is a well-understood correctness criterion that
  simplifies reasoning about the behaviour of concurrent transactions
  by ensuring they are \emph{isolated} from each other while they
  execute.  However, enforcing serializable isolation comes at a steep
  cost in performance because it necessarily restricts opportunities
  to exploit concurrency even when such opportunities would not
  violate application-specific invariants. As a result, database
  systems in practice support, and often encourage, developers to
  implement transactions using weaker alternatives. These alternatives
  break the strong isolation guarantees offered by serializable
  transactions to permit greater concurrency. Unfortunately, the
  semantics of weak isolation is poorly understood, and usually
  explained only informally in terms of low-level implementation
  artifacts. Consequently, verifying high-level correctness properties
  in such environments remains a challenging problem.

  To address this issue, we present a novel program logic that enables
  compositional reasoning about the behavior of concurrently executing
  weakly-isolated transactions.  Recognizing that the proof burden
  necessary to use this logic may dissuade application developers, we
  also describe an inference procedure based on this foundation that
  ascertains the weakest isolation level that still guarantees the
  safety of high-level consistency assertions associated with such
  transactions.  The key to effective inference is the observation
  that weakly-isolated transactions can be viewed as functional
  (monadic) computations over database state, allowing us to treat
  their operations as state transformers over the database.  This
  interpretation enables automated verification using off-the-shelf
  SMT solvers.

  Notably, our development is parametric over a transaction's specific
  isolation semantics, allowing it to be applicable over a range of
  concurrency control mechanisms.  Case studies and experiments on
  real-world applications (written in an embedded DSL in OCaml)
  demonstrate the utility of our approach, and provide strong evidence
  that automated verification of weakly-isolated transactions can be
  placed on the same formal footing as their strongly-isolated
  serializable counterparts.

\end{abstract}