### 2017-07-06 related-work.tex

\vspace*{-6pt}
\section{Related Work}
\label{sec:relatedwork}

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specifies several weak isolation levels in terms
of \emph{dependency graphs} between transactions, and the kinds of
dependencies that are forbidden in each case. The operational nature
of Adya's specifications make them suitable for runtime monitoring and
anomaly detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas
the declarative nature of our specifications make them suitable for
formal reasoning about program behaviour. Sivaramakrishnan \emph{et
al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, making it difficult
to reason about a program as it builds the trace. Cerone \emph{et
al.}~\cite{gotsmanconcur15} specify isolation levels with atomic
visibility, but their specifications are also for complete traces.
Like ours, all the aforementioned specification frameworks use the
vocabulary introduced in~\cite{burckhardt14}. However, none of them
are equipped with a reasoning framework that can use such
specifications to verify programs under weak isolation.

Recent work by Crooks \emph{et al.}~\cite{CPA+17} also explores the
use of a state-based interpretation of isolation as we do, and share
our position that enables specification of weak isolation that are not
tied to implementation-specific artifacts.  However, they do not
consider verification (manual or automated) of client programs, and it
is not immediately apparent if their specification formalism is
amenable for use within a verification toolchain.

\vspace*{-4pt}
\paragraph{Reasoning under weak isolation.} In~\cite{feketessi}, Fekete
\emph{et al.} propose a theory to characterize non-serializable
executions that arise under {\sc si}. Fekete~\cite{fekete2005} also
proposes an algorithm that allocates either {\sc si} or {\sc ser}
isolation levels to transactions while guaranteeing
serializability. In~\cite{gotsmanpodc16}, Cerone \emph{et al.} improve
on Adya's {\sc si} specification and use it to derive a static
analysis that determines the safety of substituting {\sc si} with a
weaker variant called \iso{Parallel Snapshot Isolation}~\cite{psi}.
These efforts focus on establishing the equivalence of executions
between a pair of isolation levels, without taking application
invariants into account.  Bernstein \emph{et al.}~\cite{bern2000}
propose informal semantic conditions to ensure the satisfaction of
application invariants under weaker isolation levels.  All these
techniques are tailor-made for a finite set of well-understood
isolation levels (rooted in~\cite{berenson}) under a pre-defined store
consistency model.

\vspace*{-4pt}
\paragraph{Reasoning under weak consistency.} There have been several
recent proposals to reason about programs executing under weak
consistency~\cite{bailisvldb, alvarocalm, gotsmanpopl16,redblueatc,
redblueosdi, ecinec}. All of them assume a system model that offers a
choice between a \emph{coordination-free} weak consistency level
(\emph{e.g.}, eventual consistency~\cite{redblueosdi, redblueatc,
ecinec, alvarocalm, bailisvldb}) or causal
consistency~\cite{lbc16,gotsmanpopl16}). All these efforts involve
proving that atomic and fully isolated operations preserve application
invariants when executed under these consistency levels.  In contrast,
our focus in on reasoning in the presence of weakly-isolated
transactions under a strongly consistent store.  Gotsman \emph{et
al.}~\cite{gotsmanpopl16} adapt \iso{Parallel Snapshot Isolation} to a
transaction-less setting by interpreting it as a consistency level
that serializes writes to objects; a dedicated proof rule is developed
to help prove prove program invariants hold under this model. By
parameterizing our proof system over a gamut of weak isolation
specifications, we avoid the need to define a separate proof rule for
each new isolation level we may encounter.

%% \vspace*{-4pt}
%% \paragraph{Reasoning under relaxed memory.} The reasoning mechanisms
%% used to describe and prove properties about weakly-isolated
%% transactions bear some resemblance to those used to formalize relaxed
%% memory behaviour~\cite{battycpp}.  Ridge~\cite{rgtso} generalizes
%% rely-guarantee reasoning to the x86-TSO memory model.  Likewise,
%% Vafeiadis \emph{et al.}~\cite{rsl13} generalize concurrent separation
%% logic (CSL)~\cite{csl} to the C11 relaxed memory model.  Ferreira
%% \emph{et al.}~\cite{ferreira10} propose a parameterized operational
%% semantics for relaxed memory models, but the parameterization is over
%% a relation between relaxed memory programs and related {\sc sc}
%% programs. Demange \emph{et al.}~\cite{DLZ+13} present a \emph{buffered
%%   memory model} for Java that defines an axiomatic definition for the
%% JMM in terms of memory reorderings, and an operational instantiation
%% consistent with the TSO memory model.

\vspace*{-4pt}
\paragraph{Inference.}  Prior work by Vafeiadis~\cite{Vaf10,Vaf10a}
describe \emph{action inference}, an inference procedure for computing
rely and guarantee relations in the conetxt of RGSep~\cite{VP07}, an
integration of rely-guarantee and separation logic~\cite{Rey02} that
allows one to precisely reason about local and shared state of a
concurrent program. The ideas underlying action inference have been
used to prove memory safety, linearizability, shape invariant
inference, etc.  of fine-grained concurrent data structures.  While
our motivation is similar to theirs (inference of intermediate
assertions and invariants), the context of study (transactions
vs. shared-memory concurrency), the objects being analyzed (relational
database tables vs. concurrent data structures), and the analysis
technique used to drive inference (state transformers vs. abstract
interpretation) are quite different.