### 2017-07-07 evaluation.tex

\subsection{Evaluation}

To explore the space of correct isolation levels for different transactions, we  also  ran a number  of experiments using  a real transaction OLTP workload   based on TPC-C benchmark.  
 The  workload  involves  a mix of five TPC-C transaction types: 
\C{New\_Order},  \C{Payment},  \C{Delivery}, \C{Stock Level} and  \C{Order Status}, 
operating on top of  a $10$-warehouse database.  
Each warehouse has  $10$ districts, each serving $3000$ customers. 
TPC-C  specification model  requires  maintaining multiple   consistency conditions.  We focus only on   four conditions, named $I_1$, $I_2$, $I_3$, and $I_4$.   
Consistency condition $I_1$  requires that  sales of each  warehouse  equals to the  sum of    its  district's sales.  Conditions $I_2$ and $I_3$  enforce  
 sequentially unique IDs assigned to  orders placed in the same   warehouse and district. Condition $I_4$ requires 
 the  number of  item orders   in each district 
  reflexes  the district's roll-up. 
Given the consistency conditions,  Table  \ref{tab:tpcc-eval} gives the isolation level  for the \C{New\_Order},  \C{Payment},  \C{Delivery} transactions using the  MySql and Postrgres databases   to maintain   each consistency condition. 
Recall that the \C{New\_Order} transaction places a new order for
 items from a particular district in a warehouse.  
The \C{Payment} transaction  processes  a payment for a particular customer and updates the customer, district, and warehouse tables. 
 The  \C{Delivery}  transaction processes  orders for a particular   district at  a warehouse. 
 Our experimental results  shows that  the concurrent executions of   \C{New\_Order}   
    and  \C{Delivery} transactions under a weak isolation level (i.e.,  $RC$) is anomalous, i.e., it may violate the consistency conditions $I_2$ and $I_3$. Depending on the isolation guarantees provided  by  underlying database,   the stronger  isolation level is required. For instance, the isolation level $RR$ in Postgres  is sufficient to maintain the condition while   MySql database  requires  $SER$.

 	\begin{table*}[t]\small
	\centering
\begin{tabular}{|c|c|c|c|}
		\hline
	\textbf{Transaction}   & \textbf{Invariant} 
	& \textbf{MYSQL-Isolation} & \textbf{Postgres-Isolation} \\ 
	\hline
	\multirow{4}{*}{new\_order }  & $I_1$ 
	& RC &  RC\\ 
	&  $I_2$ &SER & RR \\
	&  $I_3$ & SER  &  RR  \\
	& $I_4$ & RC & RC   \\
	\hline
	\multirow{4}{*}{Payment}  & $I_1 $ 
	& RC &  RC\\ 
	&  $I_2$  &RC & RC \\
	&  $I_3 $ & RC  &  RC  \\
	& $I_4$  & RC & RC   \\
	\hline
	\multirow{4}{*}{Delivery}  & $I_1$  
	& RC &  RC \\ 
	&  $I_2$ &SER & RR \\
	&  $I_3$ & SER  &  RR \\
	& $I_4$  & RC & RC   \\
	\hline
\end{tabular}
	\caption{ Isolation level required by TPC-C transactions. }
	\label{tab:tpcc-eval}
\end{table*}