### 2017-07-08 document.tex

While a developer can examine the isolation level for each transaction, it is 
an error-prone and difficult task to  determine  whether   the 
combination of  transaction's isolations  preserve the application correctness properties.     
Given  the three possible  isolation levels RC, RR, and SER for each five  TPC-C transaction,  
we  run  the TPC-C benchmark  under all 243  possible   isolation combinations.  Then, 
 for each combination, we  check whether  the four consistency conditions are maintained. 
Out of $243$ runs, $171$ cases fails, i.e., they could not preserve the consistency conditions.  
We have observed that \C{Delivery} and \C{New_order} transactions are  problematic when 
there are executing under a weak isolation level. 
While the   \C{New_order} transaction triggers an  invariant violation  under  weak isolation levels  (RC and RR),  
out of $160$ runs in which   the  \C{Delivery}  transaction is 
executed under  RC and RR, only $11$ runs 
fails.